### 2021-02-17 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
All files for thesis committed

On the unit interval, $[-1,1]$, we note that there is a hierarchy of orthogonal polynomials (OPs) in the sense that \cite[table 18.3.1, eqn 18.9.15]{DLMF}:
\bseqn{
	& \ddx{x} P^{(a,b)}_l (x) = \half \: (l + a + b + 1) \: P^{(a+1,b+1)}_{l-1}(x) \\
	\implies & \dmdxm{x}{m} P_l(x) = {(l+m)! \over 2^m \: l!} \: P^{(m,m)}_{l-m}(x)
}
where $P^{(a,b)}_l (x)$ is the $l$ degree \textit{Jacobi polynomial}, and $P_l(x) := P^{(0,0)}_l (x)$ is simply the \textit{Legendre polynomial} of degree $l$. Jacobi polynomials are orthogonal with weight $w(x) = (1-x)^a(1+x)^b$; that is for $l, l' \in \No,$
\bseqn{
	\int_{-1}^1 {P^{(a,b)}_l(x)}^2 \:(1-x)^a(1+x)^b \: \D x &=: \omega_{P,l}^{(a,b)} \\
	\int_{-1}^1 P^{(a,b)}_l(x) \: P^{(a,b)}_{l'}(x) \:(1-x)^a(1+x)^b \: \D x &= \omega_{P,l}^{(a,b)} \: \delta_{l, l'}.
}
Further, the \textit{associated Legendre polynomials} are a set of polynomials orthogonal with respect to unit weight on the unit interval, and are given by \bstodo{Add citation for associated legendre polys}:
\bseqn{
	\Plm(x) &:= (-1)^m (1-x^2)^\frac{m}{2} \dmdxm{x}{m} P_l(x) = \chat (1-x^2)^\frac{m}{2} P^{(m,m)}_{l-m}(x) \\
	P^{-m}_l(x) &:= \ctilde \Plm(x),
}
for $m = 0, 1, 2, \dots, l$ where
\bseqn{
	\chat &:= {(l+m)! \over (-2)^m \; l!} \\
	\ctilde &:= {(-1)^m \: (l-m)! \over (l+m)!}
}
and $n! := n \: (n-1) \: (n-2) \dots 1$ for $n \in \N$ is the standard factorial. Further, define
\bseqnnumber{
	\clm := \Bigg({(2l+1) \: (l-m)! \over 4 \pi \: (l+m)!}\Bigg)^\frac{1}{2} \:
		\begin{cases} 
			\chat \quad \quad \quad \text{if } m \ge 0 \\
			\chatmod \ctildemod \quad \text{if } m < 0
		\end{cases}. \label{eqn:clmdef}
}
This allows us to explicitly see where the normalising constants that we shall be using for our definition of the spherical harmonics come from.

Let $(x,y,z) \in \Omega$. It is useful to be able to transform between these cartesian coordinates and the spherical coordinates ($\varphi, \theta$). On this note, throughout we will use the convention that the spherical coordinate angles be defined by
\bseqn{
	x &= \sinphi \costheta = \rho(z) \costheta \\
	y &= \sinphi \sintheta = \rho(z) \sintheta \\
	z &= \cosphi.
}
We can now write down the spherical harmonics. We will use the standard definition -- that is, the spherical harmonics, orthonormal on the unit sphere, are \cite[14.30.1]{DLMF}:
\bseqnnumber{
	\Ylmfull &:= \Bigg({(2l+1) \: (l-m)! \over 4 \pi \: (l+m)!}\Bigg)^\frac{1}{2} \: \eimtheta \Plm (\cosphi) \nonumber \\
	&= \clm (1 - (\cosphi)^2)^\frac{|m|}{2} \eimtheta P^{(|m|,|m|)}_{l-|m|}(\cosphi) \nonumber \\
	&= \clm P^{(|m|,|m|)}_{l-|m|}(z) \: \rho(z)^{|m|} \: \eimtheta \label{eqn:shdef}
}
for $0 \le |m| \le l, \, l \in \No$ where $\clm$ is defined in \bsrefeqn{eqn:clmdef}, and
\bseqn{
	&\int_0^{2\pi} \int_0^\pi \Ylmfull \: {Y_{l'}^{m'}(\varphi, \theta)}^{*} \: \sinphi \: \D \varphi \: \D \theta \\
	& \quad = 2\pi \: \delta_{m, m'} \: \clm \: c_{l'}^{m'} \: \int_{-1}^1 P^{(|m|,|m|)}_{l-|m|}(z) \: P^{(|m|,|m|)}_{l'-|m|}(z) \: \rho(z)^{2|m|} \: \D z \\
	& \quad = \delta_{l, l'} \: \delta_{m, m'}.
}
where $\alpha^*$ denotes the complex conjugate of $\alpha \in \C$. Note how we can express the spherical harmonics $\Ylm$ in terms of $x,y,z$ instead of $\varphi, \theta$ by noting that $\rho(z)^{|m|} \eimtheta$ can be expressed in terms of $x,y,z$ for any $m\in\mathbb{Z}$. Indeed, they are polynomials in $x,y,z$ which we denote $\Ylm(x,y,z)$. They span all polynomials modulo the ideal generated by $x^2 + y^2 + z^2 - 1$.

@@

\begin{proof}
	The recurrence relationship for the Jacobi polynomials satisfies \cite[18.9.1, 18.9.2]{DLMF} 
\bseqn{
	z P^{(m,m)}_{l-m}(z) = \tilde \gamma_{l,m,1} P^{(m,m)}_{l-m-1}(z) + \tilde \gamma_{l,m,2} P^{(m,m)}_{l-m+1}(z),
}
for $k \ge 0, \, m \in \Z$. Hence, the expression for multiplication of the spherical harmonic $\Ylm$ by $z$ is
\bseqn{
	z\,\Ylmxyz &= \clm \eimphi \rho(z)^{|m|} z P^{(|m|,|m|)}_{l-|m|}(z)  \\
		     &= \clm \eimphi \rho(z)^{|m|} \big[ \tilde \gamma_{l,m,1} P^{(|m|,|m|)}_{l-|m|-1}(z) + \tilde \gamma_{l,m,2} P^{(|m|,|m|)}_{l-|m|+1}(z) \big] \\
		     &= \gamma_{l,m,1} Y^{m}_{l-1}(x,y,z) + \gamma_{l,m,2} Y^{m}_{l+1}(x,y,z).
}

For multiplication by $x$ and $y$, we require some further relations for the complex exponential and the Jacobi polynomials. First, recall that
\bseqn{
	\cosphi \, \eimphi &= \frac{1}{2} (e^{i\varphi} + e^{-i\varphi}) \eimphi =  \frac{1}{2} (e^{i(m+1)\varphi} + e^{i(m-1)\varphi}) \\
	\sinphi \, \eimphi &= \frac{1}{2i} (e^{i\varphi} - e^{-i\varphi}) \eimphi =  \frac{-i}{2} (e^{i(m+1)\varphi} - e^{i(m-1)\varphi})
}
Finally, we can express a Jacobi polynomial $P^{(a,b)}_{l-m}(z)$ in terms of an ultraspherical polynomial $C^{(\lambda)}_{l-m}(z)$ \cite[18.7.2]{DLMF}:
\bseqn{
	{l! \: (2m)! \over (l+m)! \: m!} \: C^{(m+1/2)}_{l-m}(z) = P^{(m,m)}_{l-m}(z),
}
where the ultraspherical polynomials satisfy the relations \cite[18.9.7, 18.9.8]{DLMF}:
\bseqn{
	C^{(\lambda)}_n(z) &= {\lambda \over n + \lambda} \: \big[ C^{(\lambda+1)}_n(z) - C^{(\lambda+1)}_{n-2}(z) \big]  \\
	(1-z^2) \, C^{(\lambda)}_n(z) &= {1 \over 4(\lambda-1)(n+\lambda)} \: \big[ (n+2\lambda-2)(n+2\lambda-1) \: C^{(\lambda-1)}_n(z) \\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad - (n+1)(n+2) \: C^{(\lambda-1)}_{n+2}(z) \big] ,
}
for $m \in \No$ where $\lambda > -\half$ is some parameter. Thus, combining these, we can write three-term recurrences for the Jacobi polynomials as:
\bseqn{
	P^{(m,m)}_{l-m}(z) &= \tilde \alpha_{l,m,4} P^{(m+1,m+1)}_{l-m}(z) + \tilde \alpha_{l,m,2} P^{(m+1,m+1)}_{l-m-2}(z) \\
	(1-z^2) \, P^{(m,m)}_{l-m}(z) &= \tilde \alpha_{l,m,3} P^{(m-1,m-1)}_{l-m+2}(z) + \tilde \alpha_{l,m,1} P^{(m-1,m-1)}_{l-m}(z),
}
for $l, m \in \No$ s.t. $0 \le m \le l$. Hence, the expressions for multiplication of the spherical harmonic $\Ylm$ by $x$ and $y$ is then:
\bseqn{
	x\,\Ylmxyz &= \clm \cosphi \: \eimphi \: \sintheta \: \rho(z)^{|m|} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= \half \clm (e^{i(m+1)\varphi} + e^{i(m-1)\varphi}) \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= \half \clm e^{i(m+1)\varphi} \rho(z)^{|m|+1} \: \big[ \tilde \alpha_{l,m,4} P^{(|m|+1,|m|+1)}_{l-|m|}(z) + \tilde \alpha_{l,m,2} P^{(|m|+1,|m|+1)}_{l-|m|-2}(z) \big] \\
	&\quad \quad + \half \clm e^{i(m-1)\varphi} \rho(z)^{|m|-1} \: \big[ \tilde \alpha_{l,m,3} P^{(|m|-1,|m|-1)}_{l-|m|+2}(z) + \tilde \alpha_{l,m,1} P^{(|m|-1,|m|-1)}_{l-|m|}(z) \big] \\
	&= \alpha_{l,m,1} Y^{m-1}_{l-1}(x,y,z) +  \alpha_{l,m,2} Y^{m+1}_{l-1}(x,y,z) \\
	&\quad \quad + \alpha_{l,m,3} Y^{m-1}_{l+1}(x,y,z) + \alpha_{l,m,4} Y^{m+1}_{l+1}(x,y,z), \\
	y\,\Ylmxyz &= \clm \sinphi \: \eimphi \: \sintheta \: \rho(z)^{|m|} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= -\half i \: \clm (e^{i(m+1)\varphi} - e^{i(m-1)\varphi}) \: \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= - i \: \big[\Alm Y^{m+1}_{l+1}(x,y,z) +  \Blm Y^{m+1}_{l-1}(x,y,z) \big] \\
	&\quad \quad \quad + i \, \big[ \Dlm Y^{m-1}_{l+1}(x,y,z) + \Elm Y^{m-1}_{l-1}(x,y,z) \big], \\
	&= i \: \Big\{ \alpha_{l,m,1} Y^{m-1}_{l-1}(x,y,z) - \alpha_{l,m,2} Y^{m+1}_{l-1}(x,y,z) \\
	&\quad \quad \quad + \alpha_{l,m,3} Y^{m-1}_{l+1}(x,y,z) - \alpha_{l,m,4} Y^{m+1}_{l+1}(x,y,z) \Big\}.
}

@@

We start as before for the scalar case by finding $x \,\gradYlm (x,y,z)$, $x \,\gradpYlm (x,y,z)$ etc. in terms of $\gradY^{m'}_{l'} (x,y,z)$, $\gradpY^{m'}_{l'} (x,y,z)$. A lovely consequence of our definitions of the scalar and vector spherical harmonics is that we can once again find explicit expressions for the coefficients in the aforementioned relations. To this end, we recall and derive some important relations of the complex exponential and Jacobi polynomials that are a part of the spherical harmonics definitions. In particular we first note that:
\bseqnnumber{
	\int_0^{2\pi} \: \eimtheta \: e^{-im'\theta} \: \cos(\theta) \D\theta &= \pi (\delta_{m',m-1} + \delta_{m',m+1}) \label{eqn:intcomplexexpcos} \\
	\int_0^{2\pi} \: \eimtheta \: e^{-im'\theta} \: \sin(\theta) \D\theta &= i \; \pi (\delta_{m',m-1} - \delta_{m',m+1}) \label{eqn:intcomplexexpsin}
}
and
\bseqnnumber{
	&\int_{-1}^1 \: \jaclmabs{l}{m}(z) \: \jaclmabs{l'}{m}(z) \: \rho(z)^{2\abs{m}} \: \D z = \delta_{l',l} \: {1 \over 2\pi (\clm)^2} \label{eqn:normjac} \\
	&\jaclmabsdz{l}{m}(z) = d_{l,m} \: \jaclmbracket{l}{\abs{m} + 1} \quad \text{where } d_{l,m} := \half (l + \abs{m} + 1), \label{eqn:diffjac},
}
where \bsrefeqn{eqn:normjac} and \bsrefeqn{eqn:normjac} are consequences of the definition of the Jacobi polynomials, see \cite[Table 18.3.1, 18.9.15]{DLMF}. One final relation we will need is that for incrementing the Jacobi parameters, which we will present in the following Lemma.
\begin{lemma}
	Let $\alpha, \beta \in \R$ be general parameters. Let $z \in [-1,1]$ and define $P_{-2}^{(\alpha+1, \beta+1)}(z) \equiv P_{-1}^{(\alpha+1, \beta+1)}(z) \equiv 0$. Then, for any $n \in \N_0$:
\bseqn{
	P_{n}^{(\alpha, \beta)}(z) =
		{1 \over 2n + \alpha + \beta + 1} \Big\{ & {(n+\alpha+\beta+1)(n+\alpha+\beta+2) \over 2n+\alpha+\beta+2} \: P_{n}^{(\alpha+1, \beta+1)}(z) \\
									  & + \big[{(n+\alpha+\beta+1)(n+\alpha) \over 2n+\alpha+\beta} \\
									  		& \quad \quad - {(n+\alpha+\beta+1)(n+\beta+1) \over 2n+\alpha+\beta+2}\big] \: P_{n-1}^{(\alpha+1, \beta+1)}(z) \\
									  & + {(n+\alpha)(n+\beta) \over 2n+\alpha+\beta} \: P_{n-2}^{(\alpha+1, \beta+1)}(z) \Big\}.
}
\end{lemma}
\begin{proof}
	This is simply as consequence of the relationships established for incrementing the parameters of Jacobi polynomials detailed in \cite[18.9.5]{DLMF}, along with the symmetry relation that $P_{n}^{(\alpha, \beta)}(-z) \equiv (-1)^n P_{n}^{(\beta, \alpha)}(z)$.
\end{proof}
We thus can present a useful corollary specific for our needs.
\begin{corollary}\label{corollary:incrementjacparams}
	Let $m \in \N_0$ and $z \in [-1,1]$. Then, for any $n \in \N_0$:
\bseqn{
	\jaclm{l}{m}(z) = \mu_{l,m,1} \: P_{l-m}^{(m+1, m+1)}(z) + \mu_{l,m,2} \: P_{l-m-2}^{(m+1, m+1)}(z) \Big],
}
where
\bseqn{
	\mu_{l,m,1} &:= {(l+m+\beta+1)(l+m+2) \over 2(2l + 1)(l+1)} \\
	\mu_{l,m,2} &:= {l \over 2(2l + 1)}.
}
\end{corollary}

The expressions for the VSHs as given in \bsrefeqn{eqn:VSHs} together with the above allow us to be able to find the coefficients for our desired expressions for $x \gradYlm$, $y \gradYlm$, $z \gradYlm$, $x \gradpYlm$, $y \gradpYlm$ and $z \gradpYlm$. 
\begin{lemma}\label{lemma:VSHrecurrences}
	For $l \in \No$, $-l \le m \le l$ the vector spherical harmonics satisfy the relationships:
	x \gradYlm &= A_{l,m,1} \gradY^{m-1}_{l-1} + A_{l,m,2} \gradY^{m+1}_{l-1} + A_{l,m,3} \gradY^{m-1}_{l+1} + A_{l,m,4} \gradY^{m+1}_{l+1} \\
& \quad \quad \quad + A_{l,m,5} \gradpY^{m-1}_{l} + A_{l,m,6} \gradpY^{m+1}_{l}, \\
	x \gradpYlm &= A_{l,m,1} \gradpY^{m-1}_{l-1} + A_{l,m,2} \gradpY^{m+1}_{l-1} + A_{l,m,3} \gradpY^{m-1}_{l+1} + A_{l,m,4} \gradpY^{m+1}_{l+1} \\
& \quad \quad \quad - A_{l,m,5} \gradY^{m-1}_{l} - A_{l,m,6} \gradY^{m+1}_{l}, \\
	y \gradYlm &= B_{l,m,1} \gradY^{m-1}_{l-1} + B_{l,m,2} \gradY^{m+1}_{l-1} + B_{l,m,3} \gradY^{m-1}_{l+1} + B_{l,m,4} \gradY^{m+1}_{l+1} \\
& \quad \quad \quad + B_{l,m,5} \gradpY^{m-1}_{l} + B_{l,m,6} \gradpY^{m+1}_{l}, \\
	y \gradpYlm &= B_{l,m,1} \gradpY^{m-1}_{l-1} + B_{l,m,2} \gradpY^{m+1}_{l-1} + B_{l,m,3} \gradpY^{m-1}_{l+1} + B_{l,m,4} \gradpY^{m+1}_{l+1} \\
& \quad \quad \quad - B_{l,m,5} \gradY^{m-1}_{l} - B_{l,m,6} \gradY^{m+1}_{l}, \\
	z \gradYlm &= \Gamma_{l,m,1} \gradY^{m}_{l-1} + \Gamma_{l,m,2} \gradY^{m}_{l+1} + \Gamma_{l,m,3} \gradpY^{m}_{l}, \\
	z \gradpYlm &= \Gamma_{l,m,1} \gradpY^{m}_{l-1} + \Gamma_{l,m,2} \gradpY^{m}_{l+1} - \Gamma_{l,m,3} \gradY^{m}_{l}
where, \bstodoinline{finish typing up (will take a while as is quite fiddly to set out)}
	A_{l,m,1} &:= \clm
		\begin{cases}

>>>>

On the unit interval, $[-1,1]$, we note that there is a hierarchy of orthogonal polynomials (OPs) in the sense that \cite[table 18.3.1, eqn 18.9.15]{DLMF}:
\bseqn{
	& \ddx{x} P^{(a,b)}_l (x) = \half \: (l + a + b + 1) \: P^{(a+1,b+1)}_{l-1}(x) \\
	\implies & \dmdxm{x}{m} P_l(x) = {(l+m)! \over 2^m \: l!} \: P^{(m,m)}_{l-m}(x)
}
where $P^{(a,b)}_l (x)$ is the $l$ degree \textit{Jacobi polynomial}, and $P_l(x) := P^{(0,0)}_l (x)$ is simply the \textit{Legendre polynomial} of degree $l$. Jacobi polynomials are orthogonal with weight $w(x) = (1-x)^a(1+x)^b$; that is for $l, l' \in \No,$
\bseqn{
	\int_{-1}^1 {P^{(a,b)}_l(x)}^2 \:(1-x)^a(1+x)^b \: \D x &=: \omega_{P,l}^{(a,b)} \\
	\int_{-1}^1 P^{(a,b)}_l(x) \: P^{(a,b)}_{l'}(x) \:(1-x)^a(1+x)^b \: \D x &= \omega_{P,l}^{(a,b)} \: \delta_{l, l'}.
}
Further, the \textit{associated Legendre polynomials} are a set of polynomials orthogonal with respect to unit weight on the unit interval, and are given by \bstodo{Add citation for associated legendre polys}:
\bseqn{
	\Plm(x) &:= (-1)^m (1-x^2)^\frac{m}{2} \dmdxm{x}{m} P_l(x) = (-1)^m \; \chat \; (1-x^2)^\frac{m}{2} \; P^{(m,m)}_{l-m}(x) \\
	P^{-m}_l(x) &:= (-1)^m \; \ctilde \; \Plm(x),
}
for $m = 0, 1, 2, \dots, l$ where
\bseqn{
	\chat &:= {(l+m)! \over 2^m \; l!} \\
	\ctilde &:= {(l-m)! \over (l+m)!}
}
and $n! := n \: (n-1) \: (n-2) \dots 1$ for $n \in \N_0$ is the standard factorial, with $0! = 1$. These relationships for the Jacobi polynomials and associated Legendre polynomials allow us to explicitly see where the normalising constants that we shall be using for our definition of the spherical harmonics come from.

On that note, we can now write down the spherical harmonics. Let $(x,y,z) \in \Omega$. It is useful to be able to transform between these cartesian coordinates and the spherical coordinates ($\varphi, \theta$). Throughout this document we will use the convention that the spherical coordinate angles be defined by
\bseqn{
	x &= \sinphi \costheta = \rho(z) \costheta \\
	y &= \sinphi \sintheta = \rho(z) \sintheta \\
	z &= \cosphi.
}
where $\rho(z) := \sqrt{1-z^2}$. We will use the standard definition -- that is, the spherical harmonics, orthonormal on the unit sphere, are \cite[14.30.1]{DLMF}:
\bseqnnumber{
	\Ylmfull &:= \Bigg({(2l+1) \: (l-m)! \over 4 \pi \: (l+m)!}\Bigg)^\frac{1}{2} \: \eimtheta \Plm (\cosphi) \nonumber \\
	&= \clm (1 - (\cosphi)^2)^\frac{|m|}{2} \eimtheta P^{(|m|,|m|)}_{l-|m|}(\cosphi) \nonumber \\
	&= \clm P^{(|m|,|m|)}_{l-|m|}(z) \: \rho(z)^{|m|} \: \eimtheta \label{eqn:shdef}
}
for $0 \le |m| \le l, \, l \in \No$, where
\bseqnnumber{
	\clm := \Bigg({(2l+1) \: (l-m)! \over 4 \pi \: (l+m)!}\Bigg)^\frac{1}{2} \:
		\begin{cases} 
			(-1)^m \: \chat &\text{if } m \ge 0 \\
			\chatmod &\text{if } m < 0
		\end{cases}. \label{eqn:clmdef}
} 
The spherical harmonics as defined are then orthonormal with respect to the inner product
\bseqn{
	&\int_0^{2\pi} \int_0^\pi \Ylmfull \: {Y_{l'}^{m'}(\varphi, \theta)}^{*} \: \sinphi \: \D \varphi \: \D \theta \\
	& \quad = 2\pi \: \delta_{m, m'} \: \clm \: c_{l'}^{m'} \: \int_{-1}^1 P^{(|m|,|m|)}_{l-|m|}(z) \: P^{(|m|,|m|)}_{l'-|m|}(z) \: \rho(z)^{2|m|} \: \D z \\
	& \quad = \delta_{l, l'} \: \delta_{m, m'}
}
where $\alpha^*$ denotes the complex conjugate of $\alpha \in \C$. Note how we can express the spherical harmonics $\Ylm$ in terms of $x,y,z$ instead of $\varphi, \theta$ by noting that $\rho(z)^{|m|} \eimtheta$ can be expressed in terms of $x,y,z$ for any $m \in \Z$. Indeed, they are polynomials in $x,y,z$ which we denote $\Ylm(x,y,z)$. They span all polynomials modulo the ideal generated by $x^2 + y^2 + z^2 - 1$.

@@

\begin{proof}
	The recurrence relationship for the Jacobi polynomials satisfies \cite[18.9.1, 18.9.2]{DLMF} 
\bseqn{
	z P^{(m,m)}_{l-m}(z) = \tilde \gamma_{l,m,1} P^{(m,m)}_{l-m-1}(z) + \tilde \gamma_{l,m,2} P^{(m,m)}_{l-m+1}(z),
}
for $k \ge 0, \, m \in \Z$. Hence, the expression for multiplication of the spherical harmonic $\Ylm$ by $z$ is
\bseqn{
	z\,\Ylmxyz &= \clm \eimtheta \rho(z)^{|m|} z P^{(|m|,|m|)}_{l-|m|}(z)  \\
		     &= \clm \eimtheta \rho(z)^{|m|} \big[ \tilde \gamma_{l,m,1} P^{(|m|,|m|)}_{l-|m|-1}(z) + \tilde \gamma_{l,m,2} P^{(|m|,|m|)}_{l-|m|+1}(z) \big] \\
		     &= \gamma_{l,m,1} Y^{m}_{l-1}(x,y,z) + \gamma_{l,m,2} Y^{m}_{l+1}(x,y,z).
}

For multiplication by $x$ and $y$, we require some further relations for the complex exponential and the Jacobi polynomials. First, recall that
\bseqn{
	\costheta \, \eimtheta &= \frac{1}{2} (e^{i\theta} + e^{-i\theta}) \eimtheta =  \frac{1}{2} (e^{i(m+1)\theta} + e^{i(m-1)\theta}) \\
	\sintheta \, \eimtheta &= \frac{1}{2i} (e^{i\theta} - e^{-i\theta}) \eimtheta =  \frac{-i}{2} (e^{i(m+1)\theta} - e^{i(m-1)\theta})
}
We can express a Jacobi polynomial $P^{(a,b)}_{l-m}(z)$ in terms of an Ultraspherical polynomial $C^{(\lambda)}_{l-m}(z)$ \cite[18.7.2]{DLMF}:
\bseqn{
	{l! \: (2m)! \over (l+m)! \: m!} \: C^{(m+1/2)}_{l-m}(z) = P^{(m,m)}_{l-m}(z),
}
where the Ultraspherical polynomials satisfy the relations \cite[18.9.7, 18.9.8]{DLMF}:
\bseqn{
	C^{(\lambda)}_n(z) &= {\lambda \over n + \lambda} \: \big[ C^{(\lambda+1)}_n(z) - C^{(\lambda+1)}_{n-2}(z) \big]  \\
	(1-z^2) \, C^{(\lambda)}_n(z) &= {1 \over 4(\lambda-1)(n+\lambda)} \: \big[ (n+2\lambda-2)(n+2\lambda-1) \: C^{(\lambda-1)}_n(z) \\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad - (n+1)(n+2) \: C^{(\lambda-1)}_{n+2}(z) \big] ,
}
for $m \in \No$ where $\lambda > -\half$ is some parameter. Thus, combining these, we can write three-term recurrences for the Jacobi polynomials as:
\bseqn{
	P^{(m,m)}_{l-m}(z) &= \tilde \alpha_{l,m,4} P^{(m+1,m+1)}_{l-m}(z) + \tilde \alpha_{l,m,2} P^{(m+1,m+1)}_{l-m-2}(z) \\
	(1-z^2) \, P^{(m,m)}_{l-m}(z) &= \tilde \alpha_{l,m,3} P^{(m-1,m-1)}_{l-m+2}(z) + \tilde \alpha_{l,m,1} P^{(m-1,m-1)}_{l-m}(z),
}
for $l, m \in \No$ s.t. $0 \le m \le l$. Hence, the expressions for multiplication of the spherical harmonic $\Ylm$ by $x$ and $y$ are then:
\bseqn{
	x \,\Ylmxyz &= \clm \costheta \: \eimtheta \: \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= \half \clm (e^{i(m+1)\theta} + e^{i(m-1)\theta}) \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= \half \clm e^{i(m+1)\theta} \rho(z)^{|m|+1} \: \big[ \tilde \alpha_{l,m,4} P^{(|m|+1,|m|+1)}_{l-|m|}(z) + \tilde \alpha_{l,m,2} P^{(|m|+1,|m|+1)}_{l-|m|-2}(z) \big] \\
	&\quad \quad + \half \clm e^{i(m-1)\theta} \rho(z)^{|m|-1} \: \big[ \tilde \alpha_{l,m,3} P^{(|m|-1,|m|-1)}_{l-|m|+2}(z) + \tilde \alpha_{l,m,1} P^{(|m|-1,|m|-1)}_{l-|m|}(z) \big] \\
	&= \alpha_{l,m,1} Y^{m-1}_{l-1}(x,y,z) +  \alpha_{l,m,2} Y^{m+1}_{l-1}(x,y,z) \\
	&\quad \quad + \alpha_{l,m,3} Y^{m-1}_{l+1}(x,y,z) + \alpha_{l,m,4} Y^{m+1}_{l+1}(x,y,z), \\
	y \,\Ylmxyz &= \clm \sintheta \: \eimtheta \: \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= -\half i \: \clm (e^{i(m+1)\theta} - e^{i(m-1)\theta}) \: \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= i \: \Big\{ \alpha_{l,m,1} Y^{m-1}_{l-1}(x,y,z) - \alpha_{l,m,2} Y^{m+1}_{l-1}(x,y,z) \\
	&\quad \quad \quad + \alpha_{l,m,3} Y^{m-1}_{l+1}(x,y,z) - \alpha_{l,m,4} Y^{m+1}_{l+1}(x,y,z) \Big\}.
}

@@

We start as before for the scalar case by finding $x \,\gradYlm (x,y,z)$, $x \,\gradpYlm (x,y,z)$ etc. in terms of $\gradY^{m'}_{l'} (x,y,z)$, $\gradpY^{m'}_{l'} (x,y,z)$. A lovely consequence of our definitions of the scalar and vector spherical harmonics is that we can once again find explicit expressions for the coefficients in the aforementioned relations. To this end, we recall and derive some important relations of the complex exponential and Jacobi polynomials that are a part of the spherical harmonics definitions. In particular we first note that:
\bseqnnumber{
	\int_0^{2\pi} \: \eimtheta \: e^{-im'\theta} \: \cos(\theta) \D\theta &= \pi (\delta_{m',m-1} + \delta_{m',m+1}) \label{eqn:intcomplexexpcos} \\
	\int_0^{2\pi} \: \eimtheta \: e^{-im'\theta} \: \sin(\theta) \D\theta &= i \; \pi (\delta_{m',m-1} - \delta_{m',m+1}) \label{eqn:intcomplexexpsin}
}
and
\bseqnnumber{
	&\int_{-1}^1 \: \jaclmabs{l}{m}(z) \: \jaclmabs{l'}{m}(z) \: \rho(z)^{2\abs{m}} \: \D z = \delta_{l',l} \: {1 \over 2\pi (\clm)^2} \label{eqn:normjac} \\
	&\jaclmabsdz{l}{m}(z) = d_{l,m} \: \jaclmbracket{l}{\abs{m} + 1} \quad \text{where } d_{l,m} := \half (l + \abs{m} + 1), \label{eqn:diffjac},
}
where \bsrefeqn{eqn:normjac} and \bsrefeqn{eqn:normjac} are consequences of the definition of the Jacobi polynomials, see \cite[Table 18.3.1, 18.9.15]{DLMF}. One final relation we will need is that for incrementing the Jacobi parameters, which we will present in the following Lemma.
\begin{lemma}
	Let $\alpha, \beta \in \R$ be general parameters. Let $z \in [-1,1]$ and define $P_{-2}^{(\alpha+1, \beta+1)}(z) \equiv P_{-1}^{(\alpha+1, \beta+1)}(z) \equiv 0$. Then, for any $n \in \N_0$:
\bseqn{
	P_{n}^{(\alpha, \beta)}(z) =
		{1 \over 2n + \alpha + \beta + 1} \Big\{ & {(n+\alpha+\beta+1)(n+\alpha+\beta+2) \over 2n+\alpha+\beta+2} \: P_{n}^{(\alpha+1, \beta+1)}(z) \\
									  & + \big[{(n+\alpha+\beta+1)(n+\alpha) \over 2n+\alpha+\beta} \\
									  		& \quad \quad - {(n+\alpha+\beta+1)(n+\beta+1) \over 2n+\alpha+\beta+2}\big] \: P_{n-1}^{(\alpha+1, \beta+1)}(z) \\
									  & + {(n+\alpha)(n+\beta) \over 2n+\alpha+\beta} \: P_{n-2}^{(\alpha+1, \beta+1)}(z) \Big\}.
}
\end{lemma}
\begin{proof}
	This is simply as consequence of the relationships established for incrementing the parameters of Jacobi polynomials detailed in \cite[18.9.5]{DLMF}, along with the symmetry relation that $P_{n}^{(\alpha, \beta)}(-z) \equiv (-1)^n P_{n}^{(\beta, \alpha)}(z)$.
\end{proof}
We thus can present a useful corollary specific for our needs.
\begin{corollary}\label{corollary:incrementjacparams}
	Let $l, m \in \N_0$ s.t. $0 \le m \le l$ and let $z \in [-1,1]$. Then:
\bseqn{
	\jaclm{l}{m}(z) = \mu_{l,m,1} \: P_{l-m}^{(m+1, m+1)}(z) + \mu_{l,m,2} \: P_{l-m-2}^{(m+1, m+1)}(z),
}
where
\bseqn{
	\mu_{l,m,1} &:= {(l+m+1)(l+m+2) \over 2(2l + 1)(l+1)} \\
	\mu_{l,m,2} &:= {l \over 2(2l + 1)}.
}
\end{corollary}
